Compositional equivalences based on Open pNets

### 2020-03-21 2016-Weak-FH-Bisim/WeakBisim.tex
appendix and page 22
Added new Figure7 in appendix, and most of the WOTs
Reviewed WOTs in page 22

$ WI_1 = \openrule
{\{\texttt{P}\mapsto \texttt{p-a}\}, [\forall \texttt{m}. \texttt{p-a} \neq \texttt{p-send(m)} ], ()}
{000 \OTWeakarrow {\tau} 000}$

$ WI_3(n) = \openrule
  {\{\texttt{P}\mapsto \texttt{p-send(m)}\}, True,
    (\texttt{s1\_msg}\gets \texttt{m}, \texttt{s1\_ec}\gets n)}
  {000 \OTWeakarrow {\underline{\texttt{in(m)}}} 100}
  \ \forall n\ge 0$

  $ WI_4(n) = \openrule
         {\{\}, True, (\texttt{m1\_msg}\gets \texttt{s1\_msg}, \texttt{m1\_ec}\gets \texttt{s1\_ec}+n, \texttt{s2\_ec}\gets \texttt{s1\_ec}+n)}
         {100 \OTWeakarrow {\tau} 210}
         \ \forall n\ge 0$

@@

  \begin{tabular}{|c|c|l|}
\hline
    Spec state & Impl state & Predicate\\
    \hline
    b0 & $000$ & True\\
    b0 & $202$ & True\\
    b1 & $100$ & $\texttt{b1\_msg = s1\_msg} \land \texttt{b1\_ec = s1\_ec}$\\
    b1 & $210$ & $\texttt{b1\_msg = m1\_msg} \land \texttt{b1\_ec = m1\_ec}$\\
    b1 & $220$ & $\texttt{b1\_msg = m2\_msg} \land \texttt{b1\_ec = m2\_ec}$\\
    b1 & $201$ & $\texttt{b1\_msg = r1\_msg} \land \texttt{b1\_ec = r1\_ec}$\\
    \hline
    \end{tabular}

@@

\begin{figure}[h]
   \centerline{\includegraphics[width=9cm]{XFIG/SimpleProtImpl-WOA}}
  \caption{Weak Open Automaton of the implementation}
   \label{Appendix:SpecOA-partial}
\end{figure}

In Figure \ref{Appendix:SpecOA-partial} we show the weak open automaton of our implementation pNet. Details of the weak transitions is listed here:

$ WI_1 = \openrule
{\{\texttt{P}\mapsto \texttt{p-a}\}, [\forall \texttt{m}. \texttt{p-a} \neq \texttt{p-send(m)} ], ()}
{000 \OTWeakarrow {\tau} 000}$

$ WI_3(n) = \openrule
  {\{\texttt{P}\mapsto \texttt{p-send(m)}\}, True,
    (\texttt{s1\_msg}\gets \texttt{m}, \texttt{s1\_ec}\gets n)}
  {000 \OTWeakarrow {\underline{\texttt{in(m)}}} 100}
  \ \forall n\ge 0$

  $ WI_4(n) = \openrule
         {\{\}, True, (\texttt{m1\_msg}\gets \texttt{s1\_msg}, \texttt{m1\_ec}\gets \texttt{s1\_ec}+n, \texttt{s2\_ec}\gets \texttt{s1\_ec}+n)}
         {100 \OTWeakarrow {\tau} 210}
         \ \forall n\ge 0$

@@

\TODO{no yet checked...:}
         $ 1 = \openrule
{\{\texttt{P}\mapsto \texttt{p-a}\}, [\forall \texttt{m}. \texttt{p-a} \neq \texttt{p-send(m)} ], ()}
{000 \OTWeakarrow {\tau} 000}$

$ 2 = \openrule
{\{\texttt{Q}\mapsto \texttt{q-b}\}, [\forall \texttt{m}. \texttt{q-b} \neq \texttt{q-recv(m)} ], ()}
{000 \OTWeakarrow {\tau} 000}$

$ 5 = \openrule
         {\{\}, True, (\texttt{m2\_ec}\gets \texttt{m1\_ec}+n)}
         {210 \OTWeakarrow {\tau} 220}
         \ \forall n\ge 0$

$ 6 = \openrule
         {\{\}, True, (\texttt{s1\_ec}\gets \texttt{m2\_ec}+n)}
         {220 \OTWeakarrow {\tau} 100}
         \ \forall n\ge 0$

$ 7 = \openrule
         {\{\}, True, (\texttt{r1\_msg}\gets \texttt{s1\_msg}, \texttt{r1\_ec}\gets \texttt{s1\_ec}+n)}
         {100 \OTWeakarrow {\tau} 201}
         \ \forall n\ge 0$

$ 7'' = \openrule
         {\{\}, True, (\texttt{r1\_msg}\gets \texttt{m1\_msg}, \texttt{r1\_ec}\gets \texttt{m2\_ec}+n)}
         {220 \OTWeakarrow {\tau} 201}
         \ \forall n\ge 1$

>>>>

$ WI_1 = \openrule
{\{\texttt{P}\mapsto \texttt{p-a}\}, [\forall \texttt{m}. \texttt{p-a} \neq \texttt{p-send(m)} ], ()}
{S \OTWeakarrow {\texttt{p-a}} S}
$ (for any state $S$)

$ WI_3(n) = \openrule
  {\{\texttt{P}\mapsto \texttt{p-send(m)}\}, True,
    (\texttt{s1\_msg}\gets \texttt{m}, \texttt{s1\_ec}\gets n)}
  {000 \OTWeakarrow {\underline{\texttt{in(m)}}} 100}
        $($\forall n\ge 0$)

$ WI_4(n) = \openrule
         {\{\}, True, 
   (\texttt{m1\_msg}\gets \texttt{s1\_msg}, \texttt{m1\_ec}\gets \texttt{s1\_ec}+n, \texttt{s2\_ec}\gets \texttt{s1\_ec}+n)}
         {100 \OTWeakarrow {\tau} 210}
        $($\forall n\ge 0$)

$ WI_{456}(n) = \openrule
         {\{\}, True, 
    (\texttt{s1\_ec}\gets \texttt{s1\_ec}+n)}
  {100 \OTWeakarrow {\tau} 100}
        $($\forall n\ge 0$)

$Post_{6a}=\ post_{4}\shortotimes\ Post_{456}^*\shortotimes\ post_{6}\\
= \left( (\texttt{m1\_msg}\gets \texttt{s1\_msg}, \texttt{m1\_ec}\gets \texttt{s1\_ec}, \texttt{s2\_ec}\gets \texttt{s1\_ec})
\shortotimes (\texttt{s1\_ec}\gets \texttt{s1\_ec}+n) \right)
\shortotimes (\texttt{s1\_ec}\gets \texttt{s2\_ec}+1) \\
= (\texttt{m1\_msg}\gets \texttt{s1\_msg}, \texttt{m1\_ec}\gets (\texttt{s2\_ec}+1)+n, \texttt{s2\_ec}\gets (\texttt{s2\_ec}+1)+n)$

$ WI_6a(n) = \openrule
         {\{\}, True, 
         (\texttt{m1\_ec}\gets \texttt{s1\_ec}+1+n, \texttt{s2\_ec}\gets \texttt{s2\_ec}+1+n)}
         {220 \OTWeakarrow {\tau} 210}
         $($\forall n\ge 0$)

$Post_{8c}= post_{8}\shortotimes\ post_{7}\shortotimes\ post_{4}\shortotimes\ Post_{456}^* 
= (\texttt{r1\_msg}\gets \texttt{s1\_msg}, \texttt{r1\_ec}\gets \texttt{s2\_msg}+1+n), \forall n\ge 0$

$ WI_{8c}(n) = \openrule
         {\{\texttt{Q}\mapsto \texttt{q-recv(s1\_msg,s2\_ec}+n)\}, True, ()}
         {210 \OTWeakarrow {\underline{\texttt{out(r1\_msg,r1\_ec)}\subst{(\texttt{r1\_msg}\gets \texttt{s1\_msg}, \texttt{r1\_ec}\gets \texttt{s2\_msg}+n)}}} 202}
        $($\forall n\ge 0$)

@@

  \begin{tabular}{|c|c|l|}
\hline
    Spec state & Impl state & Predicate\\
    \hline
    b0 & $000$ & True\\
    b0 & $202$ & True\\
    b1 & $100$ & $\texttt{b1\_msg = s1\_msg} \land \texttt{b1\_ec = s1\_ec}$\\
    b1 & $210$ & $\texttt{b1\_msg = m1\_msg} \land \texttt{b1\_ec = m1\_ec}$\\
    b1 & $220$ & $\texttt{b1\_msg = s2\_msg} \land \texttt{b1\_ec = s2\_ec}$\\
    b1 & $201$ & $\texttt{b1\_msg = r1\_msg} \land \texttt{b1\_ec = r1\_ec}$\\
    \hline
    \end{tabular}

@@

\begin{figure}[h]
   \centerline{\includegraphics[width=10cm]{XFIG/SimpleProtImpl-WOA}}
  \caption{Weak Open Automaton of the implementation}
   \label{Appendix:ImplOAl}
\end{figure}

In Figure \ref{Appendix:ImplOA} we show the weak open automaton of our implementation pNet. 
\TODO{This version is very uncomplete, at least 12 WOTs are missing, inculding: all variants of WI3 starting from 202; the inverse of WI4, WI5, WI6, WI8, etc}

\begin{figure}[h]
   \centerline{\includegraphics[width=11cm]{XFIG/SimpleProtImpl-WOA2}}
  \caption{Weak Open Automaton of the implementation: alternative}
   \label{Appendix:ImplOA2}
\end{figure}

Figure \ref{Appendix:ImplOA} shows a more compact version of this weak automation, that will be easier to read for humans...
It is based on the observation that states 202 and 000 are only linked by a "pure $\tau$" transition, and have exactly the same possible behaviors.
In this configuration we can garanty that they are Weak Bisimilar, and merge their (incoming and outgoing) transitions in the Figure.
\TODO{May be express this property more formely, but this could be also part of another paper, more oriented towards algorithms and optimisation of the Weak}

\ERIC{Correspondance between the WOTs naming:
$WI_3 ','',''' -> WI_3 a,b,c$\\
$WI_7 ','' -> WI_7 a,b$\\
$WI_8 ','',''' -> WI_8 a,b,c$\\
$WI_9 ,','' -> W_{456,564,645}$\\
$WI_{4a,5a,6a}$ are new.
}

Details of the weak transitions is listed here:

For all global states $S$:

$ WI_1 = \openrule
{\{\texttt{P}\mapsto \texttt{p-a}\}, [\forall \texttt{m}. \texttt{p-a} \neq \texttt{p-send(m)} ], ()}
{S \OTWeakarrow {\texttt{p-a}} S}$

$ WI_2 = \openrule
{\{\texttt{Q}\mapsto \texttt{q-b}\}, [\forall \texttt{m,ec}. \texttt{q-b} \neq \texttt{q-recv(m,ec)} ], ()}
{S \OTWeakarrow {\texttt{q-b}} S}$

All following transitions are parametrized by an arbitrary non-negative integer $n\in \mathcal{N}$:

$ WI_3(n) = \openrule
  {\{\texttt{P}\mapsto \texttt{p-send(m)}\}, True,
    (\texttt{s1\_msg}\gets \texttt{m}, \texttt{s1\_ec}\gets n)}
  {000 \OTWeakarrow {\underline{\texttt{in(m)}}} 100}
$

$ WI_{3a}(n) = \openrule
  {\{\texttt{P}\mapsto \texttt{p-send(m)}\}, True,
    (\texttt{m1\_msg}\gets \texttt{m}, \texttt{m1\_ec}\gets n), \texttt{s2\_ec}\gets n)}
  {000 \OTWeakarrow {\underline{\texttt{in(m)}}} 210}
$

$ WI_{3b}(n) = \openrule
  {\{\texttt{P}\mapsto \texttt{p-send(m)}\}, True,
    (\texttt{s2\_ec}\gets n)}
  {000 \OTWeakarrow {\underline{\texttt{in(m)}}} 220}
$

$ WI_{3c}(n) = \openrule
  {\{\texttt{P}\mapsto \texttt{p-send(m)}\}, True,
    (\texttt{r1\_msg}\gets \texttt{m}, \texttt{r1\_ec}\gets n)}
  {000 \OTWeakarrow {\underline{\texttt{in(m)}}} 201}
$

$ WI_4(n) = \openrule
         {\{\}, True, 
   (\texttt{m1\_msg}\gets \texttt{s1\_msg}, \texttt{m1\_ec}\gets \texttt{s1\_ec}+n, \texttt{s2\_ec}\gets \texttt{s1\_ec}+n)}
         {100 \OTWeakarrow {\tau} 210}
$

$ WI_4a(n) = \openrule
         {\{\}, True, 
    (\texttt{s2\_ec}\gets \texttt{s1\_ec}+n)}
         {100 \OTWeakarrow {\tau} 220}
$

$ WI_5(n) = \openrule
         {\{\}, True, (\texttt{s2\_ec}\gets \texttt{s2\_ec}+n)}
         {210 \OTWeakarrow {\tau} 220}
         \ \forall n\ge 0$

$ WI_5a(n) = \openrule
         {\{\}, True, (\texttt{s1\_ec}\gets \texttt{s2\_ec}+1+n)}
         {210 \OTWeakarrow {\tau} 100}
         \ \forall n\ge 0$

$ WI_6(n) = \openrule
         {\{\}, True, (\texttt{s1\_ec}\gets \texttt{s2\_ec}+1+n)}
         {220 \OTWeakarrow {\tau} 100}
         \ \forall n\ge 0$

$ WI_{456}(n) = \openrule
         {\{\}, True, 
    (\texttt{s1\_msg}\gets \texttt{s1\_msg}, \texttt{s1\_ec}\gets \texttt{s1\_ec}+n)}
  {100 \OTWeakarrow {\tau} 100}
        $($\forall n\ge 0$)

\TODO{no yet checked (old version), or not yet fully validated ...:}

$ WI_{564}(n) = \openrule
         {\{\}, True, 
    (\texttt{m1\_msg}\gets \texttt{m1\_msg}, \texttt{s2\_ec}\gets \texttt{s2\_ec}+1+n, \texttt{m1\_ec}\gets \texttt{s2\_ec}+1+n)}
  {210 \OTWeakarrow {\tau} 210}
        $($\forall n\ge 0$)

$ WI_{645}(n) = \openrule
         {\{\}, True, 
    (\texttt{s2\_ec}\gets \texttt{s2\_ec}+1+n)}
  {220 \OTWeakarrow {\tau} 220}
        $($\forall n\ge 0$)

$ WI_6a(n) = \openrule
         {\{\}, True, 
         (\texttt{m1\_ec}\gets \texttt{s1\_ec}+1+n, \texttt{s2\_ec}\gets \texttt{s2\_ec}+1+n)}
         {220 \OTWeakarrow {\tau} 210}
         \ \forall n\ge 0$

Because

$Post_{6a}=\ post_{4}\shortotimes\ Post_{456}^*\shortotimes\ post_{6}\\
= \left( (\texttt{m1\_msg}\gets \texttt{s1\_msg}, \texttt{m1\_ec}\gets \texttt{s1\_ec}, \texttt{s2\_ec}\gets \texttt{s1\_ec})
\shortotimes (\texttt{s1\_ec}\gets \texttt{s1\_ec}+n) \right)
\shortotimes (\texttt{s1\_ec}\gets \texttt{s2\_ec}+1) \\
= (\texttt{m1\_msg}\gets \texttt{s1\_msg}, \texttt{m1\_ec}\gets (\texttt{s2\_ec}+1)+n, \texttt{s2\_ec}\gets (\texttt{s2\_ec}+1)+n)$

\medskip
$ WI_7(n) = \openrule
         {\{\}, True, (\texttt{r1\_msg}\gets \texttt{s1\_msg}, \texttt{r1\_ec}\gets \texttt{s1\_ec}+n)}
         {100 \OTWeakarrow {\tau} 201}
         \ \forall n\ge 0$

$ 7'' = \openrule
         {\{\}, True, (\texttt{r1\_msg}\gets \texttt{m1\_msg}, \texttt{r1\_ec}\gets \texttt{s2\_ec}+n)}
         {220 \OTWeakarrow {\tau} 201}
         \ \forall n\ge 1$

@@
